---
title: "Patient Data Explorer"
author: "Himanshi"
date: "2026-04-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This report demonstrates basic clinical-style data analysis using a simulated patient dataset in R.

## Section 1: Data Creation

```{r}
set.seed(123)

n <- 50

patient_data <- data.frame(
  patient_id = 1:n,
  age = sample(5:70, n, replace = TRUE),
  gender = sample(c("M", "F"), n, replace = TRUE),
  treatment_grp = sample(c("D", "P"), n, replace = TRUE),
  bp = round(rnorm(n, 120, 15))
)

patient_data
```

## Section 2: Filtering Patients (Age > 30)

```{r}
filtered_data <- subset(patient_data, age > 30)
filtered_data
```

## Section 3: Blood Pressure Categorization

```{r}
filtered_data$bp_category <- ifelse(
  filtered_data$bp > 130,
  "High",
  "Normal"
)

filtered_data
```

## Section 4: Summary Statistics

```{r}
summary(filtered_data)
```

## Section 5: Mean Blood Pressure by Treatment Group

```{r}
aggregate(bp ~ treatment_grp,
          data = filtered_data,
          mean)
```

## Section 6: Age Distribution

```{r}
hist(filtered_data$age,
     col = "skyblue",
     main = "Age Distribution",
     xlab = "Age")
```

## Section 7: Blood Pressure by Treatment Group

```{r}
boxplot(bp ~ treatment_grp,
        data = filtered_data,
        col = c("orange", "green"),
        main = "Blood Pressure by Treatment Group",
        xlab = "Treatment Group",
        ylab = "Blood Pressure")
```

## Conclusion

This analysis demonstrates:
- Patient data simulation  
- Data filtering  
- Feature creation  
- Summary statistics  
- Group-wise aggregation  
- Basic data visualization